Composing like a human

### 2019-05-08 paper/main.tex
disable feature matching + better loss in GAN

% Essentials
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[ruled]{algorithm2e}

@@

\tikzset{
    neuron/.style={
        circle, draw, thick,
        inner sep=0pt,
        minimum size=3.0em,
        node distance=4em
    },
    io/.style={
        neuron,
        fill=base01
    },
    conn/.style={
        thick,
        -{Straight Barb[angle=60:2pt 3]},
    },
    connl/.style={
        thick,
        {Straight Barb[angle=60:2pt 3]}-,
    },
    prod/.style={circle, draw, thick, inner sep=0pt},
    ct/.style={circle, draw, inner sep=5pt, ultra thick,
		minimum width=10mm},
    ft/.style={circle, draw, thick, minimum width=8mm, inner sep=1pt},
    filter/.style={circle, draw, thick, minimum width=7mm, inner sep=1pt,
		path picture={
			\draw[thick, rounded corners] (path picture bounding box.center)--++(65:2mm)--++(0:1mm);
			\draw[thick, rounded corners] (path picture bounding box.center)--++(245:2mm)--++(180:1mm);}},
}

@@

        \vspace{2cm}
        Thesis Committee:\\
        Dr. Menno van Zaanen\\
        Dr. SECOND READER\\
        \vfill

@@

\chapter{The return of the Neural Net}

Over the past sixty years, Artificial Neural Networks (ANNs) have experienced a number of popularity cycles that were noted by extensive publicity of over-inflated expectations with regards to the promises of connectionism (the field of science that tries to explain mental phenomena with the help of ANNs [CITATION NEEDED]) [CITATION NEEDED]. More specifically, it promised that a system could learn to do anything it was programmed to do \parencite{minsky_perceptrons_1988}. However, each wave of praise was followed by a mass of public disappointment, criticism, and funding cuts [CITATION NEEDED]. These upturns and downturns of the field of connectionism - or more generally A.I. - are now known in the literature as the A.I. winters [CITATION NEEDED].

The recent revival of ANNs under the scope of Deep Learning (DL) has brought new ways to tackle Machine Learning problems, most notably in the context of classification (but also regression, although a less popular application; see \cite{lecun_deep_2015} for several DL examples). In addition, with the advent of chatbots \parencite{dale_return_2016}, voice-activated personal assistants \parencite{xiong_microsoft_2018}, and general-purpose A.I. systems \parencite{vinyals_starcraft_2017}, DL introduced new frontiers in artificial intelligence research.

@@

The reasons for presenting these facts will be clarified in the following sections (Section \ref{sec:thesis_goal} provides a concrete explanation) however, a terse argumentation is that progress towards general A.I. systems is marked by scant explorations of ANNs outside of the classification task, and even more so when trying to overcome some of their shortcomings, such as being highly reliant on massive amounts of data [CITATION NEEDED].
% NOTE: Maybe I was too harsh?

\section{We do not discriminate}\label{sec:no_disc}

Typically, in a neural network, the data samples go through a set of input units (that might represent pixels, word embeddings, etc.), then through multiple hidden layers, each with a given number of nodes, and reaching the output units where the answer is given \parencite{marcus_deep_2018}. An overwhelming majority of experiments with ANNs involve discriminative models \parencite{goodfellow_generative_2014}, where the goal is to assign the high-dimensional inputs to a class label \parencite[such as an animal or piece of furniture;][]{zhang_character-level_2015, krizhevsky_imagenet_2012}. However, other possible applications of DL models include (but are not limited to): regression [CITATION NEEDED], data compression \parencite{cheng_deep_2018}, and generating new data points \parencite{graves_generating_2013}.

For the purposes of this paper (detailed in Section \ref{sec:thesis_goal}), generative models are a central element in our experiments (Chapter \ref{chap:methods}). Borrowing the analogy from \textcite{goodfellow_nips_2016}, one might wonder what is the value of studying generative models. After all, in the image domain, such a framework would merely generate more images (something which the Internet as no shortage of). The answer to this lies in the fact that generative models can be used for simulating possible futures for scheduling and planning [CITATION NEEDED], providing predictions to missing data [CITATION NEEDED], or enabling work on multi-modal outputs, where a single input may have more than one correct answer \parencite{goodfellow_nips_2016}.

However, successes in generative models have been scarce, with Fully Visible Belief Networks \parencite{frey_does_1996} being the most suitable (and popular) method for the longest run. Yet, this method requires generating one sample at a time (ramping up the cost of the process to $O(n)$). Coupled with the fact that the computation is done through a neural network, creating $n$ samples requires a substantial amount of time \parencite{goodfellow_nips_2016}. Variational Autoencoders \parencite{kingma_auto-encoding_2013} are another popular method that is used for this sort of task. Their main drawback however, is that when a weak approximation of a prior or posterior distribution of the data is used, the model might learn something other than the true data distribution \parencite{goodfellow_nips_2016}.

The discovery of Generative Adversarial Networks \parencite[GANs;][]{goodfellow_generative_2014} has leapfrogged the state of developments in this field. Besides countering the drawbacks outlined above \parencite{goodfellow_nips_2016}, they have been particularly successful at a variety of tasks such as image super-resolution \parencite[; see Figure \ref{fig:srgan}]{ledig_photo-realistic_2016}, style-based image generation \parencite{karras_style-based_2018}, or image-to-image translation \parencite[such as converting a satellite image into a map, or a sketch into a photorealistic image;][;see Figure \ref{fig:pix2pix}]{isola_image--image_2016}.

In addition, although not as popular as the abovementioned, Recurrent Neural Networks (RNNs) have been successfully implemented for various generative tasks \parencite{jenal_rnn-based_2019, ICML2011Sutskever_524}. More specifically, a variation of these, the Long Short-Term Memory Networks \parencite[LSTMs;][]{hochreiter_long_1997}, are of particular interest. RNNs (and by extension LSTMs) contain a high-dimensional hidden state which allows them to remember and predict from previous inputs \parencite{ICML2011Sutskever_524}. This makes them great candidates for predicting sequences \parencite{graves_generating_2013} of text or characters, where the input at $t_0$ might be influenced by prior input at $t_{-1}$, $t_{-2}$, etc. \parencite{fan_tts_2014}. That being said, the purpose-built memory cell of an LSTM permits better storing and accessing of information \parencite{hochreiter_long_1997, graves_generating_2013} which in turn makes this RNN variation a better candidate for this sort of task.

@@

Having said all of these, the majority of developments with generative models lie in the field of Computer Vision. But that is not to say that we can generate only images. Generative models have been applied with varying degrees of success also in Natural Language Processing - where the scope was to generate high-quality text with sufficient diversity \parencite[e.g.,][]{yu_seqgan_2016, chen_adversarial_2018}. Less explored, but still successful, have been the efforts in generating music \parencite[e.g.,][]{mogren_c-rnn-gan_2016, dong_musegan_2017}. Herein lies the scope of this paper. We evaluated the samples generated by a GAN \parencite{mogren_c-rnn-gan_2016} and LSTM \parencite{oore_this_2018} when trained under a meta-learning algorithm (see Section \ref{sec:hungry}). To reiterate, a full explanation of the purpose of the paper is provided in Section \ref{sec:thesis_goal}.

\section{We always want more}\label{sec:hungry}

@@

A clear outcome from solving these issues would be that DL models are easier to train (less data) and multifunctional however, this list is not exhaustive. There is also fact that humans have the ability to generalize after one (or few) example(s) of a given object \parencite{vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016} - something which DL models tend to lack. Moreover, there are many fields where the data exhibits a large number of classes, with few examples per class. Bridging the gap between human-type learning and current learning architectures would allow models to properly capture this sort of sparse data \parencite{ravi_optimization_2016, larochelle_few-shot_2017}. This has motivated a series of explorations in the direction of "few-shot" learning (i.e., learning a class from a few examples) or more broadly, into the field of "meta-learning" \parencite[learning to learn in hopes of generalizing to new tasks;][]{chen_closer_2018, vinyals_matching_2016, zhang_metagan_2018}.

In spite of its proposed benefits, developments and experiments with few-shot learning are still scarce \parencite{larochelle_few-shot_2017}. Moreover, evaluations are largely concentrated on image data \parencite[see][]{lake_omniglot_2019, clouatre_figr_2019, vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016}. Having said that, few-shot experiments with generative networks are even less frequent; \textcite{clouatre_figr_2019}, \textcite{dong_musegan_2017} and \textcite{zhang_metagan_2018} being some of the early explorers in this direction.

@@

At the time of writing, generative models and meta-learning are two exciting areas of research in DL which have been successfully combined. However, to the knowledge of the researcher, no such experiments have been conducted in the musical domain. Thus, this paper examined the extent to which generative models can create novel and qualitative samples of music under the constraints posed by few-shot learning. In other words, it sought to evaluate the extent to which generative models can create novel and high-fidelity samples when training data is scarce. To avoid ambiguity, we define novel as new (i.e., not in the training set), and qualitative as consistent with common scales and diverse (combines chord structures with single-note sequences, is not restricted to only a few notes, etc.). This research will evaluate the generated samples of two state-of-the-art generative models (C-RNN-GAN of \cite{mogren_c-rnn-gan_2016} and Performance-RNN of \cite{oore_this_2018}) that have been trained under a meta-learning framework, Reptile \parencite{nichol_first-order_2018}, on a limited number of piano recordings. The performances of these two models were compared to the baseline proposed by \textcite{larochelle_few-shot_2017} and to recorded performances played by expert piano players at the International e-Piano Competition \parencite{university_of_minnesota_international_2019}.

@@

The following chapters will discuss more about generative models, meta-learning, and some of the models and algorithms involved in this paper (Chapter \ref{chap:theoretical_framework}), followed by a detailed description of the setup of the experiments (Chapter \ref{chap:methods}), and the results obtained from them (Chapter \ref{chap:results}). Finally, Chapter \ref{chap:discussion} closes with a discussion of the results and limitations of the experimental setup, whereas Chapter \ref{chap:conclusion} draws some conclusions that answer the above research questions.

@@

This chapter takes a deep-dive into the fields of meta-learning and generative networks in order to highlight their current state of affairs. This will lay the groundwork for the experiments detailed in Chapter \ref{chap:methods}. Section \ref{sec:rnn} will present the principal network architecture of this paper, the recurrent neural network, and argument for its use as the foundation of a generative network. Section \ref{sec:generative} will introduce Generative Adversarial Networks, a training method build specifically for the task, whereas Section \ref{sec:learn2learn} will formally introduce meta-learning, several approaches to it, and finally narrow down to current state-of-the-art algorithms.

@@

In a standard ANN, the data traverses the network through the input layer, to the hidden layer, and finally reaches the output layer; setup which is generally known as a feedforward neural network \parencite[Figure \ref{fig:feedforward_net};][]{graves_supervised_2012}. Broadly, this type of network is defined by the fact that its connections do not form cycles. If however we relax this condition, we arrive at a recurrent neural network \parencite[Figure \ref{fig:rnn_folded};][]{rumelhart_learning_1986}: a family of neural networks specifically designed for processing sequential data \parencite{goodfellow_deep_2016}. It is possible to use a feedforward architecture, specifically a one-dimensional Convolutional Neural Network, with sequential data, where the convolution would be the basis for building a time-delay neural network. Unfortunately, the output of such a network would be a function of a small number of neighbouring data points \parencite[say, the last three notes;][]{goodfellow_deep_2016} whereas in the contex of music, knowledge of the notes played from the begining of the song is needed. As such, RNNs seem to be a more compelling alternative for this task.

The information passes through the RNN in a similar manner to the feedforward network except that the activations which arrive at the hidden layer are from the current external input and prior hidden activations from the previous timestep \parencite{graves_supervised_2012}. Moreover, recurrent networks also benefit from the fact that parameters (weights) are shared across different parts of a model. This allows extending and applying the model to data points of different lengths and generalizing across all of them \parencite{goodfellow_deep_2016}. In contrast, a feedforward network would require separate parameters for each input feature, at each position. Thus, a formal definition of the network's forward pass would be
\begin{equation}
    h^t = f(h^{t-1}; x^t; \theta) \label{eq:rnn_hidden}
\end{equation}
where $h^t$ is the state of the hidden layer at time $t$, $x^t$ is the input at time $t$, and $\theta$ the model parameters.

The backward pass in a RNN is similar to that of a feedforward network: given the partial derivatives of a differentiable loss function $\mathcal{L}$ we calculate the derivatives with respect to the network's weights \parencite{graves_supervised_2012}. In this setting, a popular algorithm is called backpropagation through time \parencite[BPTT;][]{werbos_backpropagation_1990} which consists of repeated applications of the chain rule. The difference between standard backpropagation and BPTT is that the loss function depends on the activation of the hidden layer through its influence on the output layer as well as its influence on the hidden layer a the next timestep \parencite{graves_supervised_2012}.

@@

The reason why these notions have been introduced is to be able to bring up the following point. As good as standard RNNs are - in theory - at accessing contextual information when mapping input sequences to output sequences, the range of contexts that can be accessed in a standard RNN architecture is in practice quite limited \parencite{graves_supervised_2012}. \textcite{hochreiter_long_1997} outlines two problems caused by BPTT: the error signals flowing bacwards in time tend to either (1) blow up or (2) vanish. As such, the Long Short-Term Memory network has been proposed, a recurrent architecture designed to overcome these issues \parencite{hochreiter_long_1997}.

@@

Aside from being successful at a variety of classification tasks \parencite[e.g.,][]{graves_framewise_2005}, this architecture has been successfully applied to generating new data points \parencite[such as creating sequences of text;][]{graves_generating_2013}. In the context of generating musical performances, experiments with this neural architecture go as far as \textcite{eck_first_2002} however, a more noteworthy implementation is Performance-RNN \parencite{oore_this_2018}; a three layer LSTM network with 512 cells each designed to generate classical compositions whilst maintaining the playing style of expert pianists. Considering how successful it was, this model will be re-implemented in this paper's experiments (see Chapter \ref{chap:methods}).

@@

                            \draw[<-, connl] (fit.west|-int2) coordinate (aux)--++(180:7mm) node[left]{$x^t$};
                            \draw[<-, connl] ([yshift=1mm]aux)--++(135:7mm);
                            \draw[<-, connl] ([yshift=-1mm]aux)--++(-135:7mm);

                            \draw[<-, connl] (fit.north-|it) coordinate (aux)--++(90:7mm) node[above]{$x^t$};
                            \draw[<-, connl] ([xshift=1mm]aux)--++(45:7mm);
                            \draw[<-, connl] ([xshift=-1mm]aux)--++(135:7mm);

                            \draw[<-, connl] (fit.north-|ot) coordinate (aux)--++(90:7mm) node[above]{$x^t$};
                            \draw[<-, connl] ([xshift=1mm]aux)--++(45:7mm);
                            \draw[<-, connl] ([xshift=-1mm]aux)--++(135:7mm);

                            \draw[<-, connl] (fit.south-|ft) coordinate (aux)--++(-90:7mm) node[below]{$x^t$};
                            \draw[<-, connl] ([xshift=1mm]aux)--++(-45:7mm);
                            \draw[<-, connl] ([xshift=-1mm]aux)--++(-135:7mm);
    \end{tikzpicture}
    \caption{LSTM memory block with one cell. Here the gates are nonlinear summation units that collect activations from the inside and outside of the memory block and control the activation of the cell $c^t$ via multiplications. The input and output gates ($i^t$, $o^t$) multiply the block's input and output activations respectively, while the forget gate ($f^t$) multiplies the previous hidden state. The input and output activation are typically tanh or sigmoid. The only outputs ($h^t$) from the block come from the output gate multiplication. Figure adapted from \textcite{graves_supervised_2012}}
    \label{fig:lstm}
\end{figure}

@@

To make the generator's data distribution $p_g$ imitate that of the real data $p_{\text{data}}$, we need to define a prior on input noise variables $p_{z}(z)$ and then represent a mapping to the data space $G(z; \theta_g)$ where $G$ is a differentiable function represented by a network with parameters $\theta_g$. The discriminator is another network with a mapping function $D(x; \theta_d)$ that outputs a single scalar where $D(x)$ is the probability that $x$ came from the real data distribution rather than that of the generator \parencite{goodfellow_nips_2016}. We train $D$ to maximize the probability of correctly assigning a label to the training data and samples from $G$ ("real" and "fake"), while simultaneously training $G$ to minimize the dissimilarity between the two data distributions, without actually looking at $p_{\text{data}}$. In other words, $G$ must minimize $\log (1- D(G(z)))$ \parencite{goodfellow_generative_2014}.

In other words, both models play a minimax game with a value function $V(G, D)$:
\begin{equation}
    \min_G \max_D V(D, G) = \mathbb{E}_{x ~ p_{\text{data}}} [ \log D(x) ] + \mathbb{E}_{z ~ p_z (z)} [ \log (1-D(G(z))) ]
\end{equation}
where the solution is for $G$ to recover the training data distribution and $D$ to have a classification accuracy of 50\% \parencite{goodfellow_generative_2014}.

Although new, GANs are a significant conceptual and technical innovation \parencite{briot_deep_2017} which can be seen in them revitalizing interest in generative models. As mentioned before, they have been proven successful at image super-resolution \parencite{ledig_photo-realistic_2016}, image-to-image translation \parencite{isola_image--image_2016}, or style-based image generation \parencite{karras_style-based_2018}. Having said these, \textcite{mogren_c-rnn-gan_2016} has shown that GANs do not belong only to the field of Computer Vision by re-purposing this architecture to generate music. The model in question, C-RNN-GAN, features a completely unidirectional LSTM network with two layers, 350 units each as generator and a unidirectional LSTM network with a similar structure as discriminator.

@@

A summary of the parameters used in the training procedures is found in Appendix \ref{appendix:parameters}.

>>>>

% Essentials
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{lscape}
\usepackage[ruled]{algorithm2e}

@@

%% Drawing plots from PGFPlots
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepgfplotslibrary{groupplots}
\usepgfplotslibrary{dateplot}

\tikzset{
    neuron/.style={
        circle, draw, thick,
        inner sep=0pt,
        minimum size=3.0em,
        node distance=4em
    },
    io/.style={
        neuron,
        fill=base01
    },
    conn/.style={
        thick,
        -{Straight Barb[angle=60:2pt 3]},
    },
    connl/.style={
        thick,
        {Straight Barb[angle=60:2pt 3]}-,
    },
    prod/.style={circle, draw, thick, inner sep=0pt},
    ct/.style={circle, draw, inner sep=0pt, ultra thick,
		minimum size=3.0em, node distance=4em},
    ft/.style={circle, draw, thick, minimum width=8mm, inner sep=1pt},
    filter/.style={circle, draw, thick, minimum width=7mm, inner sep=1pt,
		path picture={
			\draw[thick, rounded corners] (path picture bounding box.center)--++(65:2mm)--++(0:1mm);
			\draw[thick, rounded corners] (path picture bounding box.center)--++(245:2mm)--++(180:1mm);}},
}

@@

        \vspace{2cm}
        Thesis Committee:\\
        Dr. Menno van Zaanen\\
        Dr. Martin Atzmueller\\
        \vfill

@@

\chapter{A hazy shade of winter}
%Simon and Garfunkel - A hazy shade of winter

Over the past sixty years, Artificial Neural Networks (ANNs) have experienced several popularity cycles noted in the literature by extensive publicity of over-inflated expectations with regards to the promises of connectionism (the field of science that tries to explain mental phenomena with the help of ANNs) \parencite{minsky_perceptrons_1988, knight_ai_2016, nilsson_speed_2009}. More specifically, it promised that a system could learn to do anything it was programmed to do \parencite{minsky_perceptrons_1988}. However, each wave of praise was followed by a mass of public disappointment, criticism, and funding cuts \parencite[e.g.,][]{lighthill_artificial_1972}. These upturns and downturns of the field of connectionism - or more generally A.I. - are now known in the literature as the A.I. winters \parencite{nilsson_speed_2009}.

The recent revival of ANNs under the brand of Deep Learning (DL) has brought new ways to tackle Machine Learning problems - most notably in the context of classification (but also regression, although less popular; see \cite{lecun_deep_2015} for several examples). In addition, with the advent of chatbots \parencite{dale_return_2016}, voice-activated personal assistants \parencite{xiong_microsoft_2018}, and general-purpose A.I. systems \parencite{vinyals_starcraft_2017}, DL introduced new frontiers in artificial intelligence research.

@@

The reasons for presenting these facts will be clarified bit-by-bit throughout this chapter (Section \ref{sec:thesis_goal} provides a concrete explanation) however, a terse argumentation is that progress towards general A.I. systems is marked by scant explorations of ANNs outside of the classification task, and even more so when trying to overcome some of their shortcomings, such as being highly reliant on massive amounts of data.
% NOTE: Maybe I was too harsh?

\section{Strange Fruit}\label{sec:no_disc}
% Billie Holiday - Strange Fruit

Typically, in a neural network, data samples go through a set of input units (that might represent pixels, word embeddings, etc.), then through multiple hidden layers, each with a given number of nodes, and reaching the output units where the answer is given \parencite{marcus_deep_2018}. An overwhelming majority of experiments with ANNs involve discriminative models \parencite{goodfellow_generative_2014}, where the goal is to assign high-dimensional inputs to a single class label \parencite[such as an animal or piece of furniture;][]{zhang_character-level_2015, krizhevsky_imagenet_2012}. However, other possible applications of DL models include (but are not limited to): regression \parencite{qiu_ensemble_2014}, data compression \parencite{cheng_deep_2018}, and generating new data points \parencite{graves_generating_2013}.

The last point mentioned above is particularly interesting not because it has gained a lot of traction in recent years, but because of its potential to assist the scientific community as well as the general population. Borrowing the analogy from \textcite{goodfellow_nips_2016}, one might wonder what is the value of studying generative models. If, for instance, we are dealing with images, such a framework would merely generate more images for us (something which the Internet as no shortage of). This sort of worldview, however, would only limit these systems from their full capacity; aside from the obvious use case, generative networks could also be used for simulating possible futures - such as for scheduling and planning \parencite{finn_deep_2016} -, for providing predictions to missing data, or for enabling work on multi-modal outputs, where a single input may have more than one correct answer \parencite{goodfellow_nips_2016}.

Regardless, successes in generative models have been either scarce or unattractive, with Fully Visible Belief Networks \parencite{frey_does_1996} being the most suitable (and popular) method for the longest run. Given the age and - alleged - popularity, a valid question to ask is why things haven't moved that much since then. As it turn out, this method requires generating one sample at a time (ramping up the cost of the process to $O(n)$). Coupled with the fact that the computation is done through a neural network, creating $n$ samples requires a substantial amount of time \parencite{goodfellow_nips_2016}. This discussion is not exhaustive; for a comprehensive list of various generative approaches and their downsides, refer to \textcite{goodfellow_nips_2016}.
%Variational Autoencoders \parencite{kingma_auto-encoding_2013} are another popular method that is used for this sort of task. Their main drawback however, is that when a weak approximation of a prior or posterior distribution of the data is used, the model might learn something other than the true data distribution \parencite{goodfellow_nips_2016}.

The discovery of Generative Adversarial Networks \parencite[GANs;][]{goodfellow_generative_2014} has leapfrogged the state of developments in this field. Besides countering the drawbacks outlined above, they have been particularly successful at a variety of tasks such as image super-resolution \parencite[; see Figure \ref{fig:srgan}]{ledig_photo-realistic_2016}, style-based image generation \parencite{karras_style-based_2018}, or image-to-image translation \parencite[such as converting a satellite image into a map, or a sketch into a photorealistic image;][;see Figure \ref{fig:pix2pix}]{isola_image--image_2016}.

Similarly, and although not as popular, Recurrent Neural Networks (RNNs) have been successful at various generative tasks \parencite{jenal_rnn-based_2019, ICML2011Sutskever_524}. The reason is because RNNs contain a high-dimensional hidden state which allows them to remember and predict from previous inputs \parencite{ICML2011Sutskever_524}. This made them great candidates for predicting sequences \parencite{graves_generating_2013} of text or characters, where the input at $t_0$ might be influenced by prior input at $t_{-1}$, $t_{-2}$, etc. \parencite{fan_tts_2014}. That being said, Long Short-Term Memory networks \parencite[LSTM;][]{hochreiter_long_1997} - a variant of RNNs - have been purpose-built with memory cells which permit better storing and accessing of information. This design decision variant of the recurrent network good candidates for generating sequences with long-term dependencies \parencite{graves_generating_2013}.

@@

Having said all of these, the majority of developments with generative models lie in the field of Computer Vision. But that is not to say that we can generate only images. Generative models have been applied with varying degrees of success also in Natural Language Processing - where the scope was to generate high-quality text with sufficient diversity \parencite[e.g.,][]{yu_seqgan_2016, chen_adversarial_2018}. Less explored, but still successful, have been the efforts in generating music \parencite[e.g.,][]{mogren_c-rnn-gan_2016, dong_musegan_2017}. Herein lies the scope of this paper. Given such minimal explorations into generating music, we evaluated the samples generated by a GAN \parencite{mogren_c-rnn-gan_2016} and a LSTM network \parencite{oore_this_2018} trained under the constraints imposed by few-shot learning (Section \ref{sec:hungry}) against a baseline model proposed by \textcite{larochelle_few-shot_2017}. To reiterate, a full explanation of the purpose of the paper is provided in Section \ref{sec:thesis_goal}.

\section{I want it all}\label{sec:hungry}
% Queen - I want it all

@@

A clear outcome from solving these issues would be that DL models are easier to train (less data) as well as multifunctional. Nonetheless, this list is not exhaustive. There is also fact that humans have the ability to generalize after one (or few) example(s) of a given object \parencite{vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016} - something which DL models tend to lack. Moreover, there are many fields where the data exhibits a large number of classes, with few examples per class. Bridging the gap between human-type learning and current learning architectures would allow a system to do a proper job at capturing and generalizing information from sparse data \parencite{ravi_optimization_2016, larochelle_few-shot_2017}. This has motivated a series of explorations in the direction of "few-shot" learning (i.e., learning a class from a few examples) or more broadly, into the field of "meta-learning" \parencite[learning to learn in hopes of generalizing to new tasks;][]{chen_closer_2018, vinyals_matching_2016, zhang_metagan_2018}.

In spite of its proposed benefits, developments and experiments with few-shot learning are still scarce \parencite{larochelle_few-shot_2017}. Moreover, evaluations are largely concentrated on image data \parencite[see][]{lake_omniglot_2019, clouatre_figr_2019, vinyals_matching_2016, chen_closer_2018, ravi_optimization_2016}. Having said that, few-shot experiments with generative networks are even less frequent; \textcite{clouatre_figr_2019}, \textcite{dong_musegan_2017} and \textcite{zhang_metagan_2018} being some of the early (and few) explorers in this direction.

@@

At the time of writing, generative models and meta-learning are two exciting areas of research the intersection of which could be quite promising. However, to the knowledge of the researcher, no experiments which combine the both of them have been conducted in the musical domain. Thus, this paper examined the extent to which generative models can create novel and qualitative samples of music under the constraints posed by few-shot learning. In other words, it sought to evaluate the degree to which generative models can create novel and high-fidelity samples when training data is scarce. To avoid any ambiguity, novel is used here as new (i.e., not in the training set), and qualitative as consistent with common scales and structurally diverse (combines chord structures with single-note sequences and is not restricted to only a few notes).

This research will evaluate the generated samples of two state-of-the-art generative models (C-RNN-GAN of \cite{mogren_c-rnn-gan_2016} and Performance-RNN of \cite{oore_this_2018}) that have been trained under a meta-learning framework, Reptile \parencite{nichol_first-order_2018}, with a limited number of songs. The performances of these two models were compared to the baseline proposed by \textcite{larochelle_few-shot_2017} and to recorded performances played by expert piano players at the International e-Piano Competition \parencite{university_of_minnesota_international_2019}.

@@

The following chapters will discuss more about generative models, meta-learning, and some of the models and algorithms involved in this paper (Chapter \ref{chap:theoretical_framework}), followed by a detailed description of the setup of the experiments (Chapter \ref{chap:methods}), and the results obtained from them (Chapter \ref{chap:results}). Finally, Chapter \ref{chap:discussion} closes with a discussion of the results and limitations of the experimental setup, whereas Chapter \ref{chap:conclusion} concludes the findings of this paper.

@@

This chapter takes a deep-dive into the fields of meta-learning and generative networks in order to highlight their current state of affairs. This will lay the groundwork for the experiments detailed in Chapter \ref{chap:methods}. Section \ref{sec:rnn} will present the principal network architecture of this paper, the recurrent neural network, and argument for its use as the foundation of a generative network. Section \ref{sec:generative} will introduce Generative Adversarial Networks, a training method build specifically for the task, whereas Section \ref{sec:learn2learn} will formally introduce meta-learning and some state-of-the-art approaches.

@@

In a standard ANN, the data traverses the network through the input layer, to the hidden layers, and finally the output layer; this setup is generally known as a feedforward neural network \parencite[Figure \ref{fig:feedforward_net};][]{graves_supervised_2012}. Broadly speaking, this type of network is defined by the fact that its connections do not form cycles. If however we relax this condition, we arrive at a recurrent neural network \parencite[Figure \ref{fig:rnn_folded};][]{rumelhart_learning_1986}: a family of neural networks specifically designed for processing sequential data \parencite{goodfellow_deep_2016}.

For transparency, it should be noted that it is possible to use a feedforward architecture, specifically a one-dimensional Convolutional Neural Network, with sequential data, where the convolution forms the basis of a time-delay neural network. Unfortunately, the output of such a system would be a function of a small number of neighbouring data points \parencite[such as the last and the next three notes in a song;][]{goodfellow_deep_2016}. Despite that, a system that uses convolution would quickly reach its limitation in the musical domain. Here, knowledge of the notes played from the beginning of the song is needed, meaning that we need an architecture that is able to account for long-term dependencies - which brings us back to the RNN.

The information passes through the RNN in a similar manner to the feedforward network except that the activations which arrive at the hidden layer are from the current external input and hidden activations from the previous timestep. It should be noted that the parameters of a recurrent network (weights) are shared across different timesteps \parencite{graves_supervised_2012}. This allows extending (and applying) the model to data points of different lengths whilst generalizing across all of them \parencite{goodfellow_deep_2016}. To place this into perspective, a feedforward network would require separate parameters for each input feature, at each position.

Thus, a formal definition of the network's forward pass would be
\begin{equation}
    h^t = f(h^{t-1}; x^t; \theta) \label{eq:rnn_hidden}
\end{equation}
where $h^t$ is the state of the hidden layer at time $t$, $x^t$ is the input at time $t$, and $\theta$ the model parameters.

The backward pass in a RNN is similar to that of a feedforward network: given the partial derivatives of a differentiable loss function $\mathcal{L}$ we calculate the derivatives with respect to the network's weights \parencite{graves_supervised_2012}. In this setting, a popular algorithm is called backpropagation through time \parencite[BPTT;][]{werbos_backpropagation_1990} which consists of repeated applications of the chain rule. The difference between standard backpropagation and BPTT is that the loss function depends on the activation of the hidden layer through its influence on the output layer as well as its influence on the hidden layer at the next timestep \parencite{graves_supervised_2012}.

@@

The reason why these notions have been introduced is to be able to bring up the following point. As good as standard RNNs are - in theory - at accessing contextual information when mapping input sequences to output sequences, the range of contexts that can be accessed in a standard RNN architecture is in practice quite limited \parencite{graves_supervised_2012}. \textcite{hochreiter_long_1997} outlines two problems caused by BPTT: the error signals flowing backwards in time tend to either (1) blow up or (2) vanish. As such, the Long Short-Term Memory network has been proposed, a recurrent architecture designed to overcome these issues \parencite{hochreiter_long_1997}.

@@

Aside from being successful at a variety of classification tasks \parencite[e.g.,][]{graves_framewise_2005}, this architecture has been successfully applied to generating new data points \parencite[such as creating sequences of text;][]{graves_generating_2013}. In the context of generating musical performances, experiments with this neural architecture go as far as \textcite{eck_first_2002} however, a more noteworthy implementation is Performance-RNN \parencite{oore_this_2018}; a three layer LSTM network with 512 cells each designed to generate classical compositions whilst maintaining the playing style of expert pianists. Considering its success, this model was re-implemented for the purposes of this paper's experiments (see Chapter \ref{chap:methods}).

@@

                            \draw[<-, connl] (fit.west|-int2) coordinate (aux)--++(180:7mm) node[left, io]{$x^t$};
                            \draw[<-, connl] ([yshift=1mm]aux)--++(135:7mm);
                            \draw[<-, connl] ([yshift=-1mm]aux)--++(-135:7mm);

                            \draw[<-, connl] (fit.north-|it) coordinate (aux)--++(90:7mm) node[above, io]{$x^t$};
                            \draw[<-, connl] ([xshift=1mm]aux)--++(45:7mm);
                            \draw[<-, connl] ([xshift=-1mm]aux)--++(135:7mm);

                            \draw[<-, connl] (fit.north-|ot) coordinate (aux)--++(90:7mm) node[above, io]{$x^t$};
                            \draw[<-, connl] ([xshift=1mm]aux)--++(45:7mm);
                            \draw[<-, connl] ([xshift=-1mm]aux)--++(135:7mm);

                            \draw[<-, connl] (fit.south-|ft) coordinate (aux)--++(-90:7mm) node[below, io]{$x^t$};
                            \draw[<-, connl] ([xshift=1mm]aux)--++(-45:7mm);
                            \draw[<-, connl] ([xshift=-1mm]aux)--++(-135:7mm);
    \end{tikzpicture}
    \caption{LSTM memory block with one cell. Here the gates are nonlinear summation units that collect activations from the inside and outside of the memory block and control the activation of the cell $c^t$ via multiplications. The input and output gates ($i^t$, $o^t$) multiply the block's input and output activations respectively, while the forget gate ($f^t$) multiplies the previous hidden state. The input and output activation are typically tanh or sigmoid. The only outputs ($h^t$) from the block come from the output gate multiplication. Figure adapted from \textcite{graves_supervised_2012}}
    \label{fig:lstm}
\end{figure}

@@

To make the generator's data distribution $p_g$ imitate that of the real data $p_{\text{data}}$, we need to define a prior on input noise variables $p_{z}(z)$ and then represent a mapping to the data space $G(z; \theta_g)$ where $G$ is a differentiable function represented by a network with parameters $\theta_g$. The discriminator is another network with a mapping function $D(x; \theta_d)$ that outputs a single scalar. More concretely, $D(x)$ is the probability that $x$ came from the real data distribution rather than that of the generator \parencite{goodfellow_nips_2016}. We train $D$ to maximize the probability of correctly classifying the training data and samples from $G$ (i.e., as "real" and "fake"), while simultaneously training $G$ to minimize the dissimilarity between the two data distributions, \emph{without actually looking at} $p_{\text{data}}$. In other words, $G$ must minimize $\log (1- D(G(z)))$ \parencite{goodfellow_generative_2014}.

To formalize, both models play a minimax game with a value function $V(G, D)$:
\begin{equation}
    \min_G \max_D V(D, G) = \mathbb{E}_{x ~ p_{\text{data}}} [ \log D(x) ] + \mathbb{E}_{z ~ p_z (z)} [ \log (1-D(G(z))) ]
\end{equation}
where the solution is for $G$ to recover the training data distribution and $D$ to have a classification accuracy of 50\% \parencite{goodfellow_generative_2014}.

Although new, GANs are a significant conceptual and technical innovation \parencite{briot_deep_2017} which can be seen in them revitalizing interest in generative models. As mentioned before, they have been proven successful at image super-resolution \parencite{ledig_photo-realistic_2016}, image-to-image translation \parencite{isola_image--image_2016}, or style-based image generation \parencite{karras_style-based_2018}. Having said these, \textcite{mogren_c-rnn-gan_2016} has shown that GANs do not belong only to the field of Computer Vision by successfully re-purposing this architecture to generate music. The model in question, C-RNN-GAN, features a completely unidirectional LSTM network with two layers, 350 units each as generator and a unidirectional LSTM network with a similar structure as discriminator.

@@

A summary of the parameters used in the training procedures is found in Table \ref{tab:params}.
